Visualising Life Satisfaction in Countries across Europe: Individual- and Multi-Level Evidence on Cultural and Pecuniary Factors

### 2019-01-31 LaTex/Paper_fulltext.tex
Intro and Abstract

		\begin{abstract}
			%% Text of abstract
			This project was set out to provide an easy access to a comprehensible way for exploring the differences in determinants of life satisfaction 
			all over Europe. While the actual analysis and presentation of results will be shown in different ways, this document
			shall provide an overview over the data used and the treatment of variables applied during the project. 
			The project's aim is not only to inspire interest into the determinants of happiness in human life, but also to encourage
			later researchers to critically enquire with the work done here in order to point out its shortcomings and improve the final
			result. Thus this document should provide the necessary transparency about sourcing of data and work on variables that
			has been done previously to conducting the analysis and visualising results.
		\end{abstract}

@@

	\section{Introduction and Preliminary remarks}
	Anecdotal evidence suggests that people spend a significant share of their lifetime working (for a casual calculation 
	see: \citet{thompson_what_2016}). Moreover, people's work environment is changing constantly and especially today's 
	discussion about the impact of digitalisation on the labour market requires to investigate how occupational factors shape 
	people's life in general. Since work constitutes a considerable part of people’s life, satisfaction with one’s job and the
	occupational environment must severely determine overall well-being. Drawing on the above-mentioned diversity of 
	occupations, their respective circumstances and adaptation to the ‘digital revolution’, this assumption gains even more 
	ground and relevance: \textit{If work changes, life changes.} \\
	But which work-related factors contribute to a happy - or unhappy - life? Also, how does life satisfaction of the employed
	vary across regions and countries? \\
	Using data form the European Values Study (EVS) and several other sources, we calculated multiple regression models to assess the causality of
	work-related factors and well-being on the individual level. Moreover, multi-level analyses with regard to social capital and
	macroeconomic as well as developmental indicators were performed to account for variation across regions and countries. 
	%% Könntest du den Fokus ein wenig weg von Work und hin auf multilevel + pecuniary/ non-pecuiary setzen, für so viel Fokus haben wir nicht genug work-relatedness
	\section{Motivational Manifesto}
	In a time of widespread digitalisation for many aspects of modern society, 'datafication' and subsequent analysis of data has become ever more prevalent \citep{baack_datafication_2015, lycett_datafication:_2013}. Transparency of data sourcing, data access and methods employed in the course of this analysis is a chief concern of the 'Open Science' movement \citep{delfanti_open_2010}. With more and more transparency, reproducibility of procedures is expected to increase and thus the quality of insights by replication and and improvements of methods through a 'wisdom of the (educated) crowd'. There have been extensive developments in the form of guidelines and protocols for researchers, universities and scientific journals that are aimed at promoting more transparency in the scientific process \citep{nosek_promoting_2015, miguel_promoting_2014}. 
	At the same time, large shares of the population lack the essential skills to participate in the research process or perceive the barrier of entry in acquiring the necessary skills to do so, as insurmountably high. The concept of 'Citizen Science' has thus emerged sitting at the edge of both science and education. Citizen science projects enlist non-scientist members of the public to participate in the research (mainly in data collection), allowing to survey large amounts of data, while educating participants about the research topic \citep{shirk_public_2012}. Originally coming from environmental research this approach has already allowed researcher to survey wide ranges of locations and time periods and thus provides access to high-quality data that could quite impossibly be attained by professional research teams on their own \citep{bonney_citizen_2009}.
	Conceptually, 'Open Science' and 'Citizen Science' are deeply intertwined as 'Citizen Science' needs to be as transparent and easily accessible as possible to be effective on a large scale \citep{nov_dusting_2011} and 'Open Science' needs to offer educational value and low entry barriers if it wants to gain attention and participation of a broader audience \citep{newman_future_2012}. We feel committed to the mission of these movements and have thus designed our project in order to advance it for the social sciences.

	\section{Theory and Literature Review}
	Although the causal relation between work and happiness/life satisfaction seems straightforward and intuitively correct,
	its academic investigation is far from trivial. On the one hand, this appears in light of definitional questions regarding the 
	wo variables of interest. On the other hand, operationalising the constructs appropriately remains subject to discussions
	(for subjective well-being see: \citet{kahneman_developments_2006,layard_measuring_2010}). Lastly, the causal nature of job and life
	satisfaction presents itself as far more complex than intuitively assumed. \\
	Each of these three fields is addressed by a comprehensive literature review to set the tone for the subsequent analysis.
	As this analysis draws on specific independent variables on the individual level and several factors on higher aggregate
	levels (multi-level analysis), their respective relation to life satisfaction, as assessed in the literature, is discussed as well.

>>>>

		\begin{abstract}
			%% Text of abstract
			This project was set out to provide a comprehensible way for exploring the eclectic determinants of life satisfaction 
			across Europe. An interactive visualisation tool was developed to enable independent inspections of happiness-related
			research questions. While the interactive tool covers the actual analysis and presentation of results, this document
			provides an overview over data and variables used to contribute to contemporary happiness research. 
			The project's aim is not only to spark interest on the determinants of happiness in human life, but also to encourage
			later researchers to critically reflect upon the work presented here in order to point out potential shortcomings. 
			Thus, this document should provide necessary transparency regarding data sourcing and variable selection.
			Lastly, accounts regarding the relatedness of life satisfaction and work-related variables are presented to illustrate one potential
			field of application of our interactive tool. 
		\end{abstract}

@@

	\section{Project Scope and Preliminary Remarks}
	Our project investigates life satisfaction across Europe and puts an emphasis on the data visualisation part that comes
	with communicating research results. This aspect appears most relevant with regard to `Open Science' and `Citizen
	Science' approaches that are gaining importance. In order to cater to this motivation, we developed an interactive interface
	- The European Happiness Observer - to inspect life satisfaction and its various antecedent across Europe. The app offers
	a myriad of possibilities to inspect life satisfaction on a descriptive basis or delve deeper by computing regression and multilevel
	analyses. Here, the user can select various options and combinations from a pre-selected pool of variables. \\
	Conciously, the scope of our visualisation tool is kept wider and should steer curiosity and inclination to develop own 
	happiness-related (research) questions. To assure at least some theoretical embeddedness, this paper elaborates upon the
	relation of life satisfaction and work-related variables. This perspective served as a broad orientation during the development
	phase of our tool, illustrates the variable selection and comprehensivly exhibits what kind of aspects happiness research
	can elucidate. At this point it should be stressed that the Observer enables the user to examine the questions raised in this
	paper. However, the tool's scope goes far beyond these applications and invites every user to get creative in her own 
	right. Thus, this background paper fulfills two major functions: One the one hand, it communicates the data-driven
	approach of our visualisations i.e., the European Happiness Observer app. On the other hand, it documents the theoretical foundation
	of our variable selection and examplifies a major question of happiness research, namely the relation to work-related
	variables.

	\section{Motivational Manifesto}
	In a time of widespread digitalisation for many aspects of modern society, 'datafication' and subsequent analysis of data has become ever more prevalent \citep{baack_datafication_2015, lycett_datafication:_2013}. Transparency of data sourcing, data access and methods employed in the course of this analysis is a chief concern of the 'Open Science' movement \citep{delfanti_open_2010}. With more and more transparency, reproducibility of procedures is expected to increase and thus the quality of insights by replication and and improvements of methods through a 'wisdom of the (educated) crowd'. There have been extensive developments in the form of guidelines and protocols for researchers, universities and scientific journals that are aimed at promoting more transparency in the scientific process \citep{nosek_promoting_2015, miguel_promoting_2014}. 
	At the same time, large shares of the population lack the essential skills to participate in the research process or perceive the barrier of entry in acquiring the necessary skills to do so, as insurmountably high. The concept of 'Citizen Science' has thus emerged sitting at the edge of both science and education. Citizen science projects enlist non-scientist members of the public to participate in the research (mainly in data collection), allowing to survey large amounts of data, while educating participants about the research topic \citep{shirk_public_2012}. Originally coming from environmental research this approach has already allowed researcher to survey wide ranges of locations and time periods and thus provides access to high-quality data that could quite impossibly be attained by professional research teams on their own \citep{bonney_citizen_2009}.
	Conceptually, 'Open Science' and 'Citizen Science' are deeply intertwined as 'Citizen Science' needs to be as transparent and easily accessible as possible to be effective on a large scale \citep{nov_dusting_2011} and 'Open Science' needs to offer educational value and low entry barriers if it wants to gain attention and participation of a broader audience \citep{newman_future_2012}. We feel committed to the mission of these movements and have thus designed our project in order to advance it for the social sciences.

	\section{Introduction to the Relation of Life Satisfaction and Work-Related Variables}
	Anecdotal evidence suggests that people spend a significant share of their lifetime working (for a casual calculation 
	see: \citet{thompson_what_2016}). Moreover, people's work environment is changing constantly and especially today's 
	discussion about the impact of digitalisation on the labour market requires to investigate how occupational factors shape 
	people's life in general. Since work constitutes a considerable part of people’s life, satisfaction with one’s job and the
	occupational environment must severely determine overall well-being. Drawing on the above-mentioned diversity of 
	occupations, their respective circumstances and adaptation to the ‘digital revolution’, this assumption gains even more 
	ground and relevance: \textit{If work changes, life changes.} \\
	But which work-related factors contribute to a happy - or unhappy - life? Also, how does life satisfaction of the employed
	vary across regions and countries? \\
	Using data form the European Values Study (EVS) and several other sources, we calculated multiple regression models to assess the causality of
	work-related factors and well-being on the individual level. Moreover, multi-level analyses with regard to social capital and
	macroeconomic as well as developmental indicators were performed to account for variation across regions and countries.

	\section{Theoretical Background and Methodological Implementation}
	Although the causal relation between work and happiness/life satisfaction seems straightforward and intuitively correct,
	its academic investigation is far from trivial. On the one hand, this appears in light of definitional questions regarding the 
	wo variables of interest. On the other hand, operationalising the constructs appropriately remains subject to discussions
	(for subjective well-being see: \citet{kahneman_developments_2006,layard_measuring_2010}). Lastly, the causal nature of job and life
	satisfaction presents itself as far more complex than intuitively assumed. \\
	Each of these three fields is addressed by a comprehensive literature review to set the tone for the subsequent analysis.
	As this analysis draws on specific independent variables on the individual level and several factors on higher aggregate
	levels (multi-level analysis), their respective relation to life satisfaction, as assessed in the literature, is discussed as well.
